%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Statistics for Data Science (DataSci w203)
% Week 1 homework
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Setup
\documentclass[12pt,a4paper]{article}
\usepackage[inner=1.5cm,outer=1.5cm,top=2.5cm,bottom=2.5cm]{geometry}
\usepackage{graphicx}
\usepackage[english]{babel}
\usepackage{amsmath}
\usepackage{amssymb}
\numberwithin{equation}{subsection}
\usepackage{hyperref}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


\begin{document}

\title{Statistics for Data Science \\
       Week 1 Live Session Plan}
%\institute{Institution}
\maketitle

%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
\begin{enumerate}

\item \textbf{What's statistics? What's this class?}

\item \textbf{Instructor Introduction}

\begin{enumerate}
\item Office hours Tuesdays 6:00pm - 7:00pm
\end{enumerate}

\item \textbf{Polls}

\begin{enumerate}
\item How much experience do you have with Statistics?
\item What are you most apprehensive about in this class?
\item What's the most advanced math class you've taken?
\end{enumerate}


\item \textbf{Student Introductions}
Please take just 2 minutes to tell us a tiny bit about yourself.
\begin{enumerate}
\item Your job / career interest / reason you're learning data science
\item Something very few people know about you
\end{enumerate}

\item \textbf{Weekly Workflow}\\
A typical week of the course proceeds as follows:
\begin{enumerate}
\item Before live session: Watch all async content
\item In live session: We will build upon the async to test and extend your understanding
\item After live session: complete the homework for the unit (due 24 hours before next live session)
\end{enumerate}

\item \textbf{Homework}\\
Homework is graded only with a check, check-minus, or a zero.
Each week, we will have one or two students present their homework solution from the previous unit.
Please take just 10 minutes to present your work.  A further 5 minutes will be reserved for questions.  Your homework does not need to be correct, but we expect you to be prepared to lead the class discussion, and this will count as part of your participation score.  Even though homework is not graded closely, it is the best preparation for labs and exams, so we encourage you to work hard on it.

Choose a week to be a homework presenter and sign up on our googledoc.

\item \textbf{How to Succeed in this Class}\\
Here are some strategies we think will help.
\begin{enumerate}
\item Get to know your readings, especially the Devore and Wooldridge textbooks.  (We may occasionally ask questions that you can only answer if you keep up with the readings)
\item Do the best job on the homework you can.
\item Strategize about the homework exercises with friends as much as you want, but write the final proofs and final scripts by yourself.  (However, note that you are not allowed to discusses quizzes or labs with anyone else unless otherwise specified)
\item Form study groups!
\end{enumerate}

\item \textbf{Discussion Questions}
Take 5 minutes to discuss these in breakouts, before our class-wide discussion.
\begin{enumerate}
\item Why do you need to know classical statistics, when we have fancy machine learning algorithms that can classify and predict with amazing accuracy?
\item Why isn't descriptive statistics enough for your career?
\item What mistake might result from confusing a sample mean for a population mean?
\end{enumerate}

\item \textbf{Software}
\begin{enumerate} 
    \item Apart from today I will use jupyter notebook with an R kernel to instruct you. In order so that you can participate in class you will need to install applications and packages in order to use these notebooks. 
    \item Every system is different and in the life of a data scientist figuring out how to get and install useful software packages from tutorial and question/answers on the internet is an extremely valuable skill.  
    \item Please do this by next week. 
    \item One possible way to do it. 
    \begin{itemize} 
        \item Install R from CRAN: https://cran.r-project.org/
        \item Install Anaconda: https://www.anaconda.com/download/\#macos
        \item Update conda using conda: "conda update conda"
        \item Install R-essentials package using conda: "conda install -c r r-essentials"
        \item Install any other packages from CRAN or from Anaconda Cloud
    \end{itemize}
    \item You may find R studio (https://www.rstudio.com/products/rstudio/download/ or in Anaconda) useful to complete your homework.
   
\end{enumerate}



\item \textbf{Reminders}\\
Before next week:
\begin{enumerate}
\item Complete Unit 1 homework (due 24 hours before next live session).
\item Pre Class Exercises
\item Watch all unit 2 async content.
\end{enumerate}
Lab 1 will be distributed after Week 2 live session, and you will have 2 weeks (minus 24 hours) to complete it.

\end{enumerate}

\end{document}